---
title: "Optimizing rearing of juvenile M. capitata corals"
author: "Josh Hancock, Crawford Drury (R script co author), Shayle Matsuda, Ariana S Huffmyer (R Markdown Author)"
date: "2020"
output:
  html_document:
    code_folding: hide
    toc: yes
    toc_depth: 6
    toc_float: yes
  pdf_document:
    keep_tex: yes
---

```{r, echo=TRUE, show=FALSE}
rm(list=ls(all=TRUE)) 
```

```{r setup, echo=TRUE}
knitr::opts_chunk$set(root.dir = "~/R",warning=FALSE, message=FALSE)
```

# **1. Formatting Datasets**    

Data sets are formatted to long form for growth, survivorship, and bleaching.  

In these datasets, columns are as follows:  
- "type": Aggregate or individual setter  
- "conditioned": Either conditioned (yes) or non-conditioned (no) plugs  
- "flow": Either presence of flow (yes) or low flow (no)  
- "light" & "growth_shade": Shade cloth used as high light (2x) or low light (4x)  
- "temperature": Larval conditioning temperature as ambient, cold, or high temperature  

Formatting growth.   
```{r, results=FALSE, warning=FALSE, message=FALSE}

library(readxl);library(tidyverse);library(janitor)

growth_AH<-read_xlsx("Data/Growth_03062020.xlsx")%>%clean_names()%>%
     mutate(type=case_when(str_detect(aggregate_individual, "[A-O]") ~"agg",
                           TRUE ~ "ind"))%>%
     rename(growth=growth_mm2)
```

Formatting survivorship.    
```{r, results=FALSE, warning=FALSE, message=FALSE}

raw<-read_xlsx("Data/M.cap_Survivorship_03032020.xlsx")%>%clean_names()
metadata<-raw[,1:5]

initial<-raw%>%select(plug_number,temperature,conditioned,flow,growth_shade,noquote(order(colnames(raw))))%>%
     select(-contains("part"))%>%
     select(-contains("x"))%>%
     gather(measurement,count,-plug_number,-temperature,-conditioned,-flow,-growth_shade)%>%
     separate(measurement,into=c("measurement","date"),sep=-6)%>%
     mutate(measurement=case_when(measurement=="agg_"~"agg_surv",
                                  measurement=="ind_"~"ind_surv",
                                  measurement=="agg_bleach_"~"agg_bleach",
                                  measurement=="ind_bleach_"~"ind_bleach"))%>%
     separate(measurement,into=c("type","measurement"),sep="_")%>%
     filter(date=="071519"|date=="071719")%>%
     select(-temperature,-conditioned,-flow,-growth_shade,-date)%>%
     mutate(count=as.numeric(count))%>%
     group_by(plug_number,type)%>%summarise(count=sum(count,na.rm=TRUE))%>%
     rename(initial=count)

counts<-raw%>%select(plug_number,temperature,conditioned,flow,growth_shade,noquote(order(colnames(raw))))%>%
     select(-contains("part"))%>%
     select(-contains("x"))%>%
     gather(measurement,count,-plug_number,-temperature,-conditioned,-flow,-growth_shade)%>%
     separate(measurement,into=c("measurement","date"),sep=-6)%>%
     separate(measurement,into=c("type","measurement"))%>%mutate(measurement=case_when(measurement!="bleach" ~ "surv",
                                                                                       TRUE ~ as.character(measurement)))%>%
     mutate(timepoint=case_when(date=="071519"~1, #by day from initial
                                date=="071719"~1,
                                date=="071819"~1,
                                date=="072219"~7,
                                date=="072419"~7,
                                date=="072919"~14,
                                date=="073119"~14,
                                date=="080519"~21,
                                date=="080719"~21,
                                date=="081219"~28,
                                date=="081419"~28,
                                date=="081919"~35,
                                date=="082119"~35,
                                date=="091719"~64,
                                date=="091919"~64,
                                date=="101519"~92,
                                date=="101719"~92,
                                date=="111219"~120,
                                date=="111419"~120,
                                date=="121019"~148,
                                date=="121219"~148))%>%
     select(-temperature,-conditioned,-flow,-growth_shade,-date)%>%
     mutate(count=as.numeric(count))%>%
     group_by(plug_number,timepoint,type)%>%summarise(count=sum(count,na.rm=TRUE))

surv<-left_join(left_join(counts,initial,by=c("plug_number","type"))%>%
     select(plug_number,timepoint,type,initial,count)%>%
     mutate(dead=as.numeric(initial)-as.numeric(count))%>%
     mutate(dead=case_when(dead<0 ~ 0,TRUE ~ as.numeric(dead)))%>%
     select(-initial),metadata,by="plug_number")%>%
     select(plug_number,temperature,conditioned,flow,growth_shade,timepoint,type,count,dead)%>%
     rename(alive=count)

unfold<-function(x,lost_colname,surv_colname){
     lost_expanded<-x[rep(row.names(x), x[[lost_colname]]), 1:ncol(x)]
     lost_final<-lost_expanded%>%mutate(outcome=0)
     surv_expanded<-x[rep(row.names(x), x[[surv_colname]]), 1:ncol(x)]
     surv_final<-surv_expanded %>% mutate(outcome=1)
     rbind(lost_final,surv_final)
}

surv_AH<-unfold(surv,"dead","alive") #dead=0, alive=1
```

Formatting bleaching.    
```{r, results=FALSE, warning=FALSE, message=FALSE}

bl_counts<-raw%>%select(plug_number,temperature,conditioned,flow,growth_shade,noquote(order(colnames(raw))))%>%
     select(-contains("part"))%>%
     select(-contains("x"))%>%
     gather(measurement,count,-plug_number,-temperature,-conditioned,-flow,-growth_shade)%>%
     separate(measurement,into=c("measurement","date"),sep=-6)%>%
     mutate(measurement=case_when(measurement=="agg_"~"agg_surv",
                                  measurement=="ind_"~"ind_surv",
                                  measurement=="agg_bleach_"~"agg_bleach",
                                  measurement=="ind_bleach_"~"ind_bleach"))%>%
     separate(measurement,into=c("type","measurement"),sep="_")%>%
     mutate(timepoint=case_when(date=="071519"~1,
                                date=="071719"~1,
                                date=="071819"~1,
                                date=="072219"~7,
                                date=="072419"~7,
                                date=="072919"~14,
                                date=="073119"~14,
                                date=="080519"~21,
                                date=="080719"~21,
                                date=="081219"~28,
                                date=="081419"~28,
                                date=="081919"~35,
                                date=="082119"~35,
                                date=="091719"~64,
                                date=="091919"~64,
                                date=="101519"~92,
                                date=="101719"~92,
                                date=="111219"~120,
                                date=="111419"~120,
                                date=="121019"~148,
                                date=="121219"~148))%>%
     select(-temperature,-conditioned,-flow,-growth_shade,-date)%>%
     mutate(count=as.numeric(count))%>%
     group_by(plug_number,timepoint,type,measurement)%>%summarise(count=sum(count,na.rm=TRUE))%>%
     spread(measurement,count)%>%
     mutate(total=bleach+surv)%>%
     mutate(nonbleach=total-bleach)%>%
     select(-total,-surv)


bleach<-left_join(bl_counts,metadata,by="plug_number")%>%
     select(plug_number,temperature,conditioned,flow,growth_shade,timepoint,type,bleach,nonbleach)

bleach_AH<-unfold(bleach,"bleach","nonbleach") #bleach=0, nonbleach=1

#### CLEANUP
toremove <- grep("*_AH", ls(), invert = TRUE,value = TRUE)
rm(list = c(toremove, "toremove"))
```

# **2. Growth**      

## *Growth Rate*  

Growth is calculated as total percent growth (change in growth over course of study normalized to initial size).  

Analyze growth using a linear mixed effects model.   
Fixed effects: type, flow, temperature, conditioned   
Random effects: plug number  
Subset by: 2x shade only (high light/low shade)  

First, calculate growth rates and examine data for outliers.  
```{r, results=TRUE, warning=FALSE, message=FALSE}
library(lme4); library(effects); library(ggplot2); library(car); library(lmerTest)

#calculate growth as percent change by dividing change in surface area ("growth") by the initial size ("initial_july")
growth_AH$percent_growth<-growth_AH$growth/growth_AH$initial_july
growth_AH$temperature<-as.factor(growth_AH$temperature)
growth_AH$conditioned<-as.factor(growth_AH$conditioned)
growth_AH$flow<-as.factor(growth_AH$flow)
growth_AH$light<-as.factor(growth_AH$light)
growth_AH$type<-as.factor(growth_AH$type)
growth_AH$plug_number<-as.factor(growth_AH$plug_number)

#relevel
growth_AH$temperature <- factor(growth_AH$temperature, levels = c("Cold", "Ambient", "Hot"))
growth_AH$conditioned <- factor(growth_AH$conditioned, levels = c("No", "Yes"))
growth_AH$flow <- factor(growth_AH$flow, levels = c("No", "Yes"))
growth_AH$type <- factor(growth_AH$type, levels = c("ind", "agg"))

#check distribution
hist(growth_AH$percent_growth)
max(growth_AH$percent_growth)
min(growth_AH$percent_growth)

#values within biologically expected range
```

Run full model with percent growth and fixed and random effects and examine assumptions.     

`growth_mod<-lmer(percent_growth~temperature * conditioned * flow * type + (1|plug_number), data=growth_AH)`

```{r, results=TRUE, message=FALSE, warning=TRUE}
#build model
growth_mod<-lmerTest::lmer(scale(percent_growth)~temperature * conditioned * flow * type + (1|plug_number), data=growth_AH)

#check assumptions
qqPlot(residuals(growth_mod)) #violated, there is an upper tail in the data that requires a transformation
```

The assumption of residual normality is not met, so we will conduct a square root transformation of the response variable.  

Check assumption of residual normality and homogeneity of variance.  
```{r, results=TRUE, message=FALSE, warning=TRUE}
#model with transformed data
growth_AH$growth_transformed<-sqrt(1+growth_AH$percent_growth)

growth_modT<-lmerTest::lmer(scale(growth_transformed)~temperature * conditioned * flow * type + (1|plug_number), data=growth_AH)
#square root transformation, adding constant due to negative numbers

#check assumptions
qqPlot(residuals(growth_modT)) #no longer violated
leveneTest(residuals(growth_modT)~temperature*conditioned*flow*type, data=growth_AH) #violated, use White correction in anova to account for this
```

Normality is met, but variance is violated. We will use a white correction in ANOVA summary of model to conduct conservative estimates.  

Look at summary table and ANOVA table.  
```{r, results=TRUE}
#look at summary
summary(growth_modT)
car::Anova(growth_modT, type=2, white.adjust=TRUE) #white.adjust=TRUE with car::Anova method adds heteroscedasticity-corrected adjustment for violation in variance

#output to excel file

#coefficients
x<-summary(growth_modT)$coefficients 
library(broom)
y<-broom::tidy(x)
library(writexl)
writexl::write_xlsx(as.data.frame(y), path = "Output/growth_model_coefs.xlsx")

#anova
x<-car::Anova(growth_modT, type=2, white.adjust=TRUE)
y<-broom::tidy(x)
writexl::write_xlsx(as.data.frame(y), path = "Output/growth_model_pvalues.xlsx")
```

Summarize growth with table of means and descriptive statistics.    

```{r, results=TRUE}
growth_summary <- plyr::ddply(growth_AH, c("temperature", "conditioned", "flow", "type"), summarise,
                  N    = length(percent_growth[!is.na(percent_growth)]),
                  mean = mean(percent_growth, na.rm=TRUE),
                  sd   = sd(percent_growth, na.rm=TRUE),
                  se   = sd / sqrt(N), 
                  max = max(percent_growth, na.rm=TRUE),
                  lower = mean-sd, 
                  upper = mean+sd
);growth_summary

#export to excel file
x<-growth_summary
writexl::write_xlsx(x, path = "Output/growth_descriptive_stats.xlsx")
```

## *Final Size*  

Final size is calculated as end surface area measurement for each juvenile.    

Analyze size using a linear mixed effects model.   
Fixed effects: type, flow, temperature, conditioned   
Random effects: plug number  
Subset by: 2x shade only (high light/low shade)  

First, examine distribution of total size data.  
```{r, results=TRUE, warning=FALSE, message=FALSE}
#check distribution
hist(growth_AH$final_december)
max(growth_AH$final_december)
min(growth_AH$final_december)

#values within biologically expected range
```

Run full model with total size and fixed and random effects and examine assumptions.     

`size_mod<-lmer(final_december~temperature * conditioned * flow * type + (1|plug_number), data=growth_AH)`

```{r, results=TRUE, message=FALSE, warning=TRUE}
#build model
size_mod<-lmerTest::lmer(scale(final_december)~temperature * conditioned * flow * type + (1|plug_number), data=growth_AH)

#check assumptions
qqPlot(residuals(size_mod)) #violated, there is an upper tail in the data that requires a transformation
```

The assumption of residual normality is not met, so we will conduct a 1/10 transformation of the response variable.  

Check assumption of residual normality and homogeneity of variance.  
```{r, results=TRUE, message=FALSE, warning=TRUE}
#model with transformed data
growth_AH$size_transformed<-(growth_AH$final_december)^(1/10)

size_modT<-lmerTest::lmer(scale(size_transformed)~temperature * conditioned * flow * type + (1|plug_number), data=growth_AH)
#square root transformation

#check assumptions
qqPlot(residuals(size_modT)) #no longer violated
leveneTest(residuals(size_modT)~temperature*conditioned*flow*type, data=growth_AH) #violated, use White correction in anova to account for this
hist(residuals(size_modT))
```

Normality is met, but variance is violated. We will use a white correction in ANOVA summary of model to conduct conservative estimates.  

Look at summary table and ANOVA table.  
```{r, results=TRUE}
#look at summary
summary(size_modT)
car::Anova(size_modT, type=2, white.adjust=TRUE) #white.adjust=TRUE with car::Anova method adds heteroscedasticity-corrected adjustment for violation in variance

#output to excel file

#coefficients
x<-summary(size_modT)$coefficients 
library(broom)
y<-broom::tidy(x)
library(writexl)
writexl::write_xlsx(as.data.frame(y), path = "Output/size_model_coefs.xlsx")

#anova
x<-car::Anova(size_modT, type=2, white.adjust=TRUE)
y<-broom::tidy(x)
writexl::write_xlsx(as.data.frame(y), path = "Output/size_model_pvalues.xlsx")
```

Summarize growth with table of means and descriptive statistics.    

```{r, results=TRUE}
size_summary <- plyr::ddply(growth_AH, c("temperature", "conditioned", "flow", "type"), summarise,
                  N    = length(final_december[!is.na(final_december)]),
                  mean = mean(final_december, na.rm=TRUE),
                  sd   = sd(final_december, na.rm=TRUE),
                  se   = sd / sqrt(N), 
                  max = max(final_december, na.rm=TRUE),
                  lower = mean-sd, 
                  upper = mean+sd
);size_summary

#export to excel file
x<-size_summary
writexl::write_xlsx(x, path = "Output/size_descriptive_stats.xlsx")
```



# **3. Survivorship**  

Survivorship is measured as a binomial response (1 = alive, 0 = dead) at the final timepoint of the experiment.  

Format survivorship data frame.  
```{r, results=TRUE}
#format data
surv_AH$temperature<-as.factor(surv_AH$temperature)
surv_AH$conditioned<-as.factor(surv_AH$conditioned)
surv_AH$flow<-as.factor(surv_AH$flow)
surv_AH$growth_shade<-as.factor(surv_AH$growth_shade)
surv_AH$type<-as.factor(surv_AH$type)
surv_AH$plug_number<-as.factor(surv_AH$plug_number)
surv_AH$timepoint<-as.factor(surv_AH$timepoint)
```

### *Surivorship by shade*  

Run survivorship model for shade only to characterize complete mortality in high shade treatment.  
Analyze with a binomial mixed effect model.    
Fixed effects: shade, timepoint
Random effects: plug number  

`surv_shade_mod<-glmer(outcome ~ growth_shade * timepoint + (1|plug_number), data=surv_AH, family=binomial)`  

Run model and check for overdispersion.  
```{r, results=TRUE, message=FALSE, warning=FALSE}
library(blmeco)

surv_shade_mod<-glmer(outcome ~ growth_shade * timepoint + (1|plug_number), data=surv_AH, family=binomial)

dispersion_glmer(surv_shade_mod) #no evidence of overdispersion
```

View summary and ANOVA table.  

```{r, results=TRUE}
summary(surv_shade_mod)
Anova(surv_shade_mod)

#export to excel
#coefficients
x<-summary(surv_shade_mod)$coefficients 
y<-broom::tidy(x)
writexl::write_xlsx(as.data.frame(y), path = "Output/surv_shade_model_coefs.xlsx")

#anova
x<-car::Anova(surv_shade_mod, type=2)
y<-broom::tidy(x)
writexl::write_xlsx(as.data.frame(y), path = "Output/surv_shade_model_pvalues.xlsx")
```

Summarize and display statistics for survival over time by shade treatment.   
```{r, results=TRUE}

surv_shade_summary <- plyr::ddply(surv_AH, c("timepoint", "growth_shade"), summarise,
                  N    = length(outcome[!is.na(outcome)]),
                  mean = mean(outcome, na.rm=TRUE),
                  sd   = sd(outcome, na.rm=TRUE),
                  se   = sd / sqrt(N), 
                  max = max(outcome, na.rm=TRUE),
                  lower = mean-sd, 
                  upper = mean+sd
);surv_shade_summary

#export to excel file
x<-surv_shade_summary
writexl::write_xlsx(x, path = "Output/surv_shade_descriptive_stats.xlsx")
```

Conduct post-hoc test for timepoint.  

```{r, results=TRUE}
library(emmeans); library(multcomp)
emm = emmeans(surv_shade_mod, ~ growth_shade | timepoint) #can choose other adjustment methods
pairs(emm)
emm_results<-cld(emm, Letters=c(LETTERS)) #letter display

#export to excel file
x<-emm_results
writexl::write_xlsx(x, path = "Output/surv_shade_posthoc.xlsx")
```

### *Survivorship within low shade*  

Analyze with a binomial mixed effect model.    
Fixed effects: temperature, flow, conditioned, type  
Random effects: plug number  
Subset: final timepoint, growth_shade = X2  

`surv_mod<-glmer(outcome ~ temperature * flow * conditioned * type + (1|plug_number), data=surv_AH, family=binomial, subset=c(growth_shade=="X2", timepoint=="148"))`  

Run model and check for overdispersion.  
```{r, results=TRUE, message=FALSE, warning=FALSE}

#run model in only 2x shade and type as random effect and final timepoint 
surv_mod<-glmer(outcome ~ temperature * flow * conditioned * type + (1|plug_number), data=surv_AH, family=binomial, subset=c(growth_shade=="X2", timepoint=="148"))

dispersion_glmer(surv_mod) #no evidence of overdispersion
```

View summary and ANOVA table.  

```{r, results=TRUE}
summary(surv_mod)
Anova(surv_mod)

#export to excel
#coefficients
x<-summary(surv_mod)$coefficients 
y<-broom::tidy(x)
writexl::write_xlsx(as.data.frame(y), path = "Output/surv_model_coefs.xlsx")

#anova
x<-car::Anova(surv_mod, type=2)
y<-broom::tidy(x)
writexl::write_xlsx(as.data.frame(y), path = "Output/surv_model_pvalues.xlsx")
```

Summarize and display statistics for survival at the final timepoint within X2 shade treatment.   
```{r, results=TRUE}
finalsurv<-surv_AH[which(surv_AH$timepoint == "148"),]
finalsurv2<-finalsurv[which(finalsurv$growth_shade == "X2"),]

surv_summary <- plyr::ddply(finalsurv2, c("temperature", "conditioned", "flow", "type"), summarise,
                  N    = length(outcome[!is.na(outcome)]),
                  mean = mean(outcome, na.rm=TRUE),
                  sd   = sd(outcome, na.rm=TRUE),
                  se   = sd / sqrt(N), 
                  max = max(outcome, na.rm=TRUE),
                  lower = mean-sd, 
                  upper = mean+sd
);surv_summary

#export to excel file
x<-surv_summary
writexl::write_xlsx(x, path = "Output/surv_descriptive_stats.xlsx")
```


# **4. Bleaching**  

Bleaching is measured as a binary response over the course of the study (0=bleached, 1=nonbleached).  

Analyze with a binomial mixed effects model.  
Fixed effects: temperature, flow, conditioned, type 
Random effects: timepoint (days), plug number
Subsets: growth_shade = X2 (low shade) 

Format bleaching dataset.  
```{r, results=TRUE}
#format data
bleach_AH$temperature<-as.factor(bleach_AH$temperature)
bleach_AH$conditioned<-as.factor(bleach_AH$conditioned)
bleach_AH$flow<-as.factor(bleach_AH$flow)
bleach_AH$growth_shade<-as.factor(bleach_AH$growth_shade)
bleach_AH$type<-as.factor(bleach_AH$type)
bleach_AH$plug_number<-as.factor(bleach_AH$plug_number)
bleach_AH$timepoint<-as.factor(bleach_AH$timepoint)
```

Analyze model.  

`bleach_mod<-glmer(outcome ~ temperature * flow * conditioned * type + (1|timepoint) + (1|plug_number), data=bleach_AH, family = binomial, subset=c(growth_shade=="X2"))`

Check for overdispersion.  
```{r, results=TRUE, warning=FALSE, message=FALSE}
bleach_mod<-glmer(outcome ~ temperature * flow * conditioned * type + (1|timepoint) + (1|plug_number), data=bleach_AH, family = binomial, subset=c(growth_shade=="X2")) #timepoint is number of days

dispersion_glmer(bleach_mod)
```

View summary and anova table. 

````{r, results=TRUE}
summary(bleach_mod)
Anova(bleach_mod)

#export to excel
#coefficients
x<-summary(bleach_mod)$coefficients 
y<-broom::tidy(x)
writexl::write_xlsx(as.data.frame(y), path = "Output/bleach_model_coefs.xlsx")

#anova
x<-car::Anova(bleach_mod, type=2)
y<-broom::tidy(x)
writexl::write_xlsx(as.data.frame(y), path = "Output/bleach_model_pvalues.xlsx")
``` 

Generate a table of bleaching at each timepoint.  
```{r, results=TRUE}
sumbleach<-bleach_AH[which(bleach_AH$growth_shade == "X2"),]

bleach_summary <- plyr::ddply(sumbleach, c("timepoint", "temperature", "conditioned", "flow", "type"), summarise,
                  N    = length(outcome[!is.na(outcome)]),
                  mean = mean(outcome, na.rm=TRUE),
                  sd   = sd(outcome, na.rm=TRUE),
                  se   = sd / sqrt(N), 
                  max = max(outcome, na.rm=TRUE),
                  lower = mean-sd, 
                  upper = mean+sd
);bleach_summary

#export to excel file
x<-bleach_summary
writexl::write_xlsx(x, path = "Output/bleach_descriptive_stats.xlsx")

```


# **5. Effect Sizes**  

Calculate Cohen's d effect sizes for main effects of juvenile coral responses.  

Calculate effect sizes for growth and set reference levels for controls.  

Controls: 
Type = Individual
Flow = No Flow
Conditioning = Not Conditioned
Shade = Low Shade
Temperature = Ambient

Using a Cohen's d Welch test, which does not assume equal variance. Hedge's correction not needed due to large sample size.  
```{r, results=TRUE}
library(effsize); library(rstatix)
growth_AH$type <- relevel(growth_AH$type, ref = "agg")
growth_AH$flow <- relevel(growth_AH$flow, ref = "Yes")
growth_AH$conditioned <- relevel(growth_AH$conditioned, ref = "Yes")

cohens_d(percent_growth ~ type, var.equal = FALSE, ref.group="agg", data=growth_AH)
cohens_d(percent_growth ~ flow, var.equal = FALSE, ref.group="Yes", data=growth_AH)
cohens_d(percent_growth ~ conditioned, var.equal = FALSE, ref.group="Yes", data=growth_AH)

#subset temperature to remove levels and compare between ambient and cool and ambient and high
#compare hot vs ambient
growth_AH$temperature <- relevel(growth_AH$temperature, ref = "Hot")
growth_AH_h<-growth_AH[!(growth_AH$temperature=="Cold"),] #remove cold
growth_AH_h$temperature <- factor(growth_AH_h$temperature)
cohens_d(percent_growth ~ temperature, var.equal = FALSE, ref.group="Hot", data=growth_AH_h)

#compare cold vs ambient
growth_AH$temperature <- relevel(growth_AH$temperature, ref = "Cold")
growth_AH_c<-growth_AH[!(growth_AH$temperature=="Hot"),] #remove cold
growth_AH_c$temperature <- factor(growth_AH_c$temperature)
cohens_d(percent_growth ~ temperature, var.equal = FALSE, ref.group="Cold", data=growth_AH_c)

```

Calculate effect sizes for survivorship. 

```{r, results=TRUE}

finalsurv$type <- relevel(finalsurv$type, ref = "agg")
finalsurv$flow <- relevel(finalsurv$flow, ref = "Yes")
finalsurv$conditioned <- relevel(finalsurv$conditioned, ref = "Yes")
finalsurv$growth_shade <- relevel(finalsurv$growth_shade, ref = "X4")

cohens_d(outcome ~ type, var.equal = FALSE, ref.group="agg", data=finalsurv)
cohens_d(outcome ~ flow, var.equal = FALSE, ref.group="Yes", data=finalsurv)
cohens_d(outcome ~ conditioned, var.equal = FALSE, ref.group="Yes", data=finalsurv)

#subset temperature to remove levels and compare between ambient and cool and ambient and high
#compare hot vs ambient
finalsurv$temperature <- relevel(finalsurv$temperature, ref = "Hot")
finalsurv_h<-finalsurv[!(finalsurv$temperature=="Cold"),] #remove cold
finalsurv_h$temperature <- factor(finalsurv_h$temperature)
cohens_d(outcome ~ temperature, var.equal = FALSE, ref.group="Hot", data=finalsurv_h)

#compare cold vs ambient
finalsurv$temperature <- relevel(finalsurv$temperature, ref = "Cold")
finalsurv_c<-finalsurv[!(finalsurv$temperature=="Hot"),] #remove cold
finalsurv_c$temperature <- factor(finalsurv_c$temperature)
cohens_d(outcome ~ temperature, var.equal = FALSE, ref.group="Cold", data=finalsurv_c)
```

Calculate effect sizes for bleaching.   

```{r, results=TRUE}
bleach_AH<-as.data.frame(bleach_AH)
bleach_AH$type <- relevel(bleach_AH$type, ref = "agg")
bleach_AH$flow <- relevel(bleach_AH$flow, ref = "Yes")
bleach_AH$conditioned <- relevel(bleach_AH$conditioned, ref = "Yes")
bleach_AH$growth_shade <- relevel(bleach_AH$growth_shade, ref = "X4")

cohens_d(outcome ~ type, var.equal = FALSE, ref.group="agg", data=bleach_AH)
cohens_d(outcome ~ flow, var.equal = FALSE, ref.group="Yes", data=bleach_AH)
cohens_d(outcome ~ conditioned, var.equal = FALSE, ref.group="Yes", data=bleach_AH)
cohens_d(outcome ~ growth_shade, var.equal = FALSE, ref.group="X2", data=bleach_AH)

#subset temperature to remove levels and compare between ambient and cool and ambient and high
#compare hot vs ambient
bleach_AH$temperature <- relevel(bleach_AH$temperature, ref = "Hot")
bleach_AH_h<-bleach_AH[!(bleach_AH$temperature=="Cold"),] #remove cold
bleach_AH_h$temperature <- factor(bleach_AH_h$temperature)
cohens_d(outcome ~ temperature, var.equal = FALSE, ref.group="Hot", data=bleach_AH_h)

#compare cold vs ambient
bleach_AH$temperature <- relevel(bleach_AH$temperature, ref = "Cold")
bleach_AH_c<-bleach_AH[!(bleach_AH$temperature=="Hot"),] #remove cold
bleach_AH_c$temperature <- factor(bleach_AH_c$temperature)
cohens_d(outcome ~ temperature, var.equal = FALSE, ref.group="Cold", data=bleach_AH_c)
```


# **Correlation of Growth and Survival**  

To analyze the relationship between growth and survival for individuals and aggregates on each plug.   

```{r, results=TRUE}
#summarize growth by plug for individuals and aggregates
growth_trt <- plyr::ddply(growth_AH, c("type", "flow", "conditioned", "temperature"), summarise,
                  growth = mean(percent_growth, na.rm=TRUE)
)

#summarize survivorship by plug for individuals and aggregates
surv_trt <- plyr::ddply(finalsurv, c("type", "flow", "conditioned", "temperature"), summarise,
                  survival = mean(outcome, na.rm=TRUE)
)

#create new dataframe and match values to include plug, type, growth, and survival 

reg_data<-merge(growth_trt, surv_trt)
```

View correlation for individuals and aggregates separately. 

```{r, results=TRUE}
reg_data_agg<-reg_data%>%filter(type=="agg")
cor.test(reg_data_agg$survival, reg_data_agg$growth, method=c("spearman"))

reg_data_ind<-reg_data%>%filter(type=="ind")
cor.test(reg_data_ind$survival, reg_data_ind$growth, method=c("spearman"))
```



```{r, results=TRUE, warning=FALSE}
Reg1<-ggplot(reg_data, aes(x=growth, y=survival)) + 
  geom_point(aes(color=type), size=3) +
  geom_smooth(method='lm', aes(color = type), fullrange=TRUE, se=TRUE)+
   scale_color_manual(name="Type",
                    values=c("blue", "black"), 
                    labels=c("Aggregate", "Individual"))+
  theme_classic()+
  ylim(0,1)+
  theme(text = element_text(size = 18))+
  theme(axis.text = element_text(size = 18, color="black"))+
  theme(axis.title = element_text(size = 18, face="bold"))+
  theme(legend.title = element_text(size = 18, face="bold"))+
  theme(legend.text = element_text(size = 18, color="black"))+
  ylab(expression(bold(paste("Survival")))) +
  xlab(expression(bold(paste("Growth Rate")))); Reg1

ggsave("Output/Surv_growth_regression_Figure.pdf", plot=Reg1, height=8, width=10, units = c("in"), dpi=300) #output figure

```